Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems

### 2024-06-15 final/oup-authoring-template.tex
Remove centering
Added ORCID
Change to thin figures

\author[1,2]{Mauro Silberberg\ORCID{0000-0000-0000-0000}}
\author[3]{Henning Hermjakob\ORCID{0000-0000-0000-0000}}
\author[3]{Rahuman S. Malik-Sheriff\ORCID{0000-0000-0000-0000}}
\author[1,2,$\ast$]{Hernán E. Grecco\ORCID{0000-0000-0000-0000}}

@@

@@

\subsection{Extensible definition of reaction networks usingSimBio}
\label{extensible-definition-of-reaction-networks-using-simbio}

@@

%% sample for biography with author's image
% \begin{biography}{{\color{black!20}\rule{77pt}{77pt}}}{\author{Author Name.} This is sample author biography text. The values provided in the optional argument are meant for sample purposes. There is no need to include the width and height of an image in the optional argument for live articles. This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
% \end{biography}

%% sample for biography without author's image
% \begin{biography}{}{\author{Author Name.} This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
% \end{biography}

% FIGURE 1
\begin{figure*}[t]
  \begin{minipage}[t]{\linewidth}

  {\centering

  \begin{CodeInput}
  \begin{Highlighting}[]
  \KeywordTok{from} \ClassTok{poincare} \KeywordTok{import} \ClassTok{Derivative}, \ClassTok{System}, \ClassTok{Variable}, \FunctionTok{initial}
  \end{Highlighting}
  \end{CodeInput}

  }

  \end{minipage}%
  \newline
  \begin{minipage}[t]{\linewidth}

  {\centering

  \begin{minipage}[c]{0.6\linewidth}

  {\centering

  \begin{CodeInput}
  \begin{Highlighting}[]
  \KeywordTok{class} \ClassTok{Decay}\KeywordTok{(}\ClassTok{System}\KeywordTok{)}\OperatorTok{:}
      \VariableTok{x}\OperatorTok{:} \ClassTok{Variable} \OperatorTok{=} \FunctionTok{initial}\KeywordTok{(\VariableTok{default}\OperatorTok{=}\ValueTok{1})}
      \VariableTok{eq} \OperatorTok{=} \VariableTok{x}.\FunctionTok{derive}\KeywordTok{()} \OperatorTok{\textless{}\textless{}} \OperatorTok{-}\VariableTok{x}
  \end{Highlighting}
  \end{CodeInput}

  }

  \end{minipage}%
  %
  \begin{minipage}[c]{0.2\linewidth}

  {\centering

  \[
  \frac{dx}{dt} = -x
  \]

  }

  \end{minipage}%
  %
  \begin{minipage}[c]{0.2\linewidth}

  {\centering

  \[
  x(0) = 1
  \]

  }

  \end{minipage}%

  \subcaption{\label{fig-first-order}First-order system of an exponential decay. The variable
  \texttt{x} stores the initial condition for \(x\), and the variable
  \texttt{eq} stores the rate equation for \(x\).}

  }

  \end{minipage}%
  \newline
  \begin{minipage}[t]{\linewidth}

  {\centering

  \begin{minipage}[c]{0.6\linewidth}

  {\centering

  \begin{CodeInput}
  \begin{Highlighting}[]
  \KeywordTok{class} \ClassTok{Oscillator}\KeywordTok{(}\ClassTok{System}\KeywordTok{)}\OperatorTok{:}
      \VariableTok{x}: \ClassTok{Variable} \OperatorTok{=} \FunctionTok{initial}\KeywordTok{(\VariableTok{default}\OperatorTok{=}\ValueTok{1})}
      \VariableTok{v}: \ClassTok{Derivative} \OperatorTok{=} \VariableTok{x}.\FunctionTok{derive}\KeywordTok{(\VariableTok{initial}\OperatorTok{=}\ValueTok{0})}
      \VariableTok{eq} \OperatorTok{=} \VariableTok{v}.\FunctionTok{derive}\KeywordTok{()} \OperatorTok{\textless{}\textless{}} \OperatorTok{-}\VariableTok{x}
  \end{Highlighting}
  \end{CodeInput}

  }

  \end{minipage}%
  %
  \begin{minipage}[c]{0.2\linewidth}

  {\centering

  \[
  \frac{d^2x}{dt^2} = -x
  \]

  }

  \end{minipage}%
  %
  \begin{minipage}[c]{0.2\linewidth}

  {\centering

  \[
  \begin{cases}
      x(0) &= 1 \\
      \frac{dx}{dt}(0) &= 0
  \end{cases}
  \]

  }

  \end{minipage}%

  \subcaption{\label{fig-second-order}Second-order system of an harmonic oscillator, where the
  variable \texttt{v} stores the derivative \(\frac{dx}{dt}\) and the rate
  equation is specified for the derivative \texttt{v}.}

  }

  \end{minipage}%
  \newline
  \begin{minipage}[t]{\linewidth}

  {\centering

  \begin{minipage}[c]{0.6\linewidth}

  {\centering

  \begin{CodeInput}
  \begin{Highlighting}[]
  \KeywordTok{class} \ClassTok{BigModel}\KeywordTok{(}\ClassTok{System}\KeywordTok{)}:
      \VariableTok{x}: \ClassTok{Variable} \OperatorTok{=} \FunctionTok{initial}\KeywordTok{(\VariableTok{default}\OperatorTok{=}\ValueTok{1})}
      \VariableTok{linked} \OperatorTok{=} \ClassTok{Decay}\KeywordTok{(\VariableTok{x}\OperatorTok{=}\VariableTok{x})}
      \VariableTok{independent} \OperatorTok{=} \ClassTok{Decay}\KeywordTok{(\VariableTok{x}\OperatorTok{=}\ValueTok{2})}
  \end{Highlighting}
  \end{CodeInput}

  }

  \end{minipage}%
  %
  \begin{minipage}[c]{0.2\linewidth}

  {\centering

  \[
  \begin{cases}
      \frac{dx}{dt} = -x \\
      \frac{dx_{ind}}{dt} = -x_{ind}
  \end{cases}
  \]

  }

  \end{minipage}%
  %
  \begin{minipage}[c]{0.2\linewidth}

  {\centering

  \[
  \begin{cases}
      x(0) &= 1 \\
      x_{ind}(0) &= 2
  \end{cases}
  \]

  }

  \end{minipage}%

  \subcaption{\label{fig-composition}First-order system of two exponential decays by composition of
  the \texttt{Decay} system. The subsystem \texttt{linked} has a reference
  to the outer variable \texttt{x}, while the subsystem
  \texttt{independent} defines a new variable with initial condition $2$,
  which on the corresponding mathematical expression was named \(x_{ind}\).}

  }

  \end{minipage}%

  \end{figure*}

% FIGURE 2
\begin{figure*}[t]

  \begin{minipage}[b]{0.50\linewidth}
  {
      \centering 
      \includegraphics[width=\textwidth]{ide1.pdf}
  }
  \end{minipage}%
  %
  \begin{minipage}[b]{0.50\linewidth}
  {
      \centering
      \includegraphics[width=\textwidth]{ide2.pdf}
  }
  \end{minipage}%

  \caption{
      \label{fig-ide}
      Screenshots of Visual Studio Code showing
      tooltips (green arrows) and
      highlighted type errors (red arrows).
      On the left,
      we show that \texttt{a},
      a \texttt{Constant} assigned with \texttt{assign(...,\ constant=True)},
      can be used for \texttt{Variable b}'s initial condition.
      Instead,
      it is flagged as a type error (red underlining)
      when using \texttt{c}, a \texttt{Parameter},
      for \texttt{Variable d}'s initial condition,
      The IDE automatically recognizes \texttt{e} as an \texttt{Equation},
      and provides autocompletion of the \texttt{Model}'s components.
      A tooltip is shown when composing models (green arrow, right),
      which show the expected variables and their default values.
      The IDE highlights wrong names (\texttt{z} is not a name in \texttt{Model})
      and mismatched types (\texttt{x} is \texttt{Variable} and \texttt{a} must be a number or a \texttt{Constant})
  }

  \end{figure*}

% FIGURE 3
\begin{figure*}[t]

  \begin{minipage}[t]{\linewidth}

  {\centering

  \begin{CodeInput}
  \begin{Highlighting}[]
  \KeywordTok{import} \ClassTok{numpy} \KeywordTok{as} \ClassTok{np}
  \KeywordTok{from} \ClassTok{poincare} \KeywordTok{import} \ClassTok{Simulator}

  \VariableTok{sim} \OperatorTok{=} \ClassTok{Simulator}\KeywordTok{(\ClassTok{Oscillator})}
  \VariableTok{df} \OperatorTok{=} \VariableTok{sim}.\FunctionTok{solve}\KeywordTok{(}\VariableTok{save\_at}\OperatorTok{=}\ClassTok{np}.\FunctionTok{linspace}\KeywordTok{(}\ValueTok{0}, \ValueTok{5}, \ValueTok{101}\KeywordTok{))}
  \end{Highlighting}
  \end{CodeInput}

  }

  \end{minipage}%
  \newline
  \begin{minipage}[t]{0.45\linewidth}

  {\centering

  \begin{CodeInput}
  \begin{Highlighting}[]
  \VariableTok{df}.\FunctionTok{head}\KeywordTok{()}
  \end{Highlighting}
  \end{CodeInput}

  \begin{CodeOutput}
      \begin{tabular}[]{@{}lll@{}}
          \toprule
          & x & v \\
          time & & \\
          \midrule
          0.00 & 1.000000 & 0.000000 \\
          0.05 & 0.998750 & -0.049979 \\
          0.10 & 0.995005 & -0.099835 \\
          0.15 & 0.988772 & -0.149440 \\
          0.20 & 0.980069 & -0.198672 \\
          \botrule
      \end{tabular}
  \end{CodeOutput}

  }

  \end{minipage}%
  \hfill
  \begin{minipage}[t]{0.45\linewidth}

  {\centering

  \begin{CodeInput}
  \begin{Highlighting}[]
  \VariableTok{df}.\FunctionTok{plot}\KeywordTok{()}\OperatorTok{;}
  \end{Highlighting}
  \end{CodeInput}

  \begin{CodeOutput}
    \includegraphics[width=\textwidth]{cell-output}
  \end{CodeOutput}

  }

  \end{minipage}%

  \caption{\label{fig-sim}Simulation of the \texttt{Oscillator} system
  from Figure~\ref{fig-second-order}. The output is a
  \texttt{pandas.DataFrame} with a column for each variable and the time
  as index. It is inspected and plotted with the \texttt{pandas} methods.}

  \end{figure*}

% FIGURE 4
\begin{figure*}[t]

  \begin{minipage}[t]{\linewidth}
    \centering 
    \begin{CodeInput}
    \begin{Highlighting}[]
    \KeywordTok{from}\ClassTok{ simbio }\KeywordTok{import}\ClassTok{ Compartment, MassAction, Species, RateLaw, \FunctionTok{initial}}
    \end{Highlighting}
    \end{CodeInput}
  \end{minipage}%
  \newline
  \begin{minipage}[t]{\linewidth}
    \centering 
    \begin{minipage}[c]{0.60\linewidth}
      \centering 
      \begin{CodeInput}
      \begin{Highlighting}[]
        \KeywordTok{class}\ClassTok{ Model}\KeywordTok{(}\ClassTok{Compartment}\KeywordTok{)}:
            \CommentTok{"""2A {-}\textgreater{} B"""}

        \VariableTok{    A}: \ClassTok{Species }\OperatorTok{=}\FunctionTok{ initial}\KeywordTok{(}\VariableTok{default}\OperatorTok{=}\ValueTok{1}\KeywordTok{)}
        \VariableTok{    B}: \ClassTok{Species }\OperatorTok{=}\FunctionTok{ initial}\KeywordTok{(}\VariableTok{default}\OperatorTok{=}\ValueTok{0}\KeywordTok{)}
        \VariableTok{    r }\OperatorTok{=}\ClassTok{ RateLaw}\KeywordTok{(}
        \VariableTok{        reactants}\OperatorTok{=}\KeywordTok{[}\ValueTok{2} \OperatorTok{*}\VariableTok{ A}\KeywordTok{]},
        \VariableTok{        products}\OperatorTok{=}\NormalTok{\KeywordTok{[}B\KeywordTok{]},}
        \VariableTok{        rate}\OperatorTok{=}\ValueTok{1}\NormalTok{,}
        \KeywordTok{    )}
      \end{Highlighting}
      \end{CodeInput}
    \end{minipage}%
    %
    \begin{minipage}[c]{0.40\linewidth}
      \centering 
      \[
      \begin{cases}
          \frac{dA}{dt} = -2 \\
          \frac{dB}{dt} = +1
      \end{cases}
      \begin{cases}
          A(0) &= 1 \\
          B(0) &= 0
      \end{cases}
      \]
    \end{minipage}%
  \end{minipage}%
  \newline
  \begin{minipage}[t]{\linewidth}
      \centering 
      \begin{minipage}[c]{0.60\linewidth}
          \centering 
          \begin{CodeInput}
          \begin{Highlighting}[]
          \KeywordTok{class}\ClassTok{ Model}\KeywordTok{(}\ClassTok{Compartment}\KeywordTok{)}:
              \CommentTok{"""2A {-}\textgreater{} B"""}

          \VariableTok{    A}: \ClassTok{Species }\OperatorTok{=}\FunctionTok{ initial}\KeywordTok{(}\VariableTok{default}\OperatorTok{=}\ValueTok{1}\KeywordTok{)}
          \VariableTok{    B}: \ClassTok{Species }\OperatorTok{=}\FunctionTok{ initial}\KeywordTok{(}\VariableTok{default}\OperatorTok{=}\ValueTok{0}\KeywordTok{)}
          \VariableTok{    r }\OperatorTok{=}\ClassTok{ MassAction}\KeywordTok{(}
          \VariableTok{        reactants}\OperatorTok{=}\KeywordTok{[}\ValueTok{2} \OperatorTok{*}\VariableTok{ A}\KeywordTok{]},
          \VariableTok{        products}\OperatorTok{=}\NormalTok{\KeywordTok{[}B\KeywordTok{]},}
          \VariableTok{        rate}\OperatorTok{=}\ValueTok{1}\NormalTok{,}
          \KeywordTok{    )}
          \end{Highlighting}
          \end{CodeInput}
      \end{minipage}%
      %
      \begin{minipage}[c]{0.40\linewidth}
          \centering 
          \[
          \begin{cases}
              \frac{dA}{dt} = -2 A^2 \\
              \frac{dB}{dt} = +1 A^2
          \end{cases}
          \begin{cases}
              A(0) &= 1 \\
              B(0) &= 0
          \end{cases}
          \]
      \end{minipage}%
  \end{minipage}%

  \caption{
    A reaction system for species \(A\) and \(B\)
    with initial conditions \(1\) and \(0\), respectively. A single reaction
    transforming \(2A\) into \(B\) is saved in variable \texttt{r}. The rate
    \(1\) is specified directly for \texttt{RateLaw}, and is proportional to
    the reactants for \texttt{MassAction}.
  }
  \label{fig-simbio}
\end{figure*}

% FIGURE 5
\begin{figure*}[t]
  \centering

  \begin{CodeInput}
  \begin{Highlighting}[]
  \KeywordTok{from}\ClassTok{ simbio.io }\KeywordTok{import}\ClassTok{ biomodels, sbml}

  \VariableTok{model }\OperatorTok{=}\ClassTok{ sbml}.\FunctionTok{load}\KeywordTok{(}\StringTok{"repressilator.sbml"}\KeywordTok{)}  \CommentTok{\# from existing SBML file}
  \VariableTok{model }\OperatorTok{=}\ClassTok{ biomodels}.\FunctionTok{load}\KeywordTok{(}\StringTok{"BIOMD12"}\KeywordTok{)}  \CommentTok{\# from BioModels}
  \VariableTok{model}
  \end{Highlighting}
  \end{CodeInput}

  \begin{CodeOutput}
      \begin{tabular}{@{}lll@{}}
          \multicolumn{3}{l}{Elowitz2000 - Repressilator} \\
          \toprule
          type & total & names \\
          \midrule
          variables  &  6 & PX, PY, PZ, X, Y, Z \\
          parameters & 17 & cell, beta, alpha0, alpha, eff, n, KM, tau\_mRNA, tau\_prot, ... \\
          equations  & 12 & Reaction1, Reaction2, Reaction3, Reaction4, Reaction5, ... \\
          \botrule
      \end{tabular}
  \end{CodeOutput}

  \caption{Creation of a model from a local SBML file or one uploaded to BioModels.}
  \label{fig-simbio-io}
\end{figure*}

>>>>

\author[1,2]{Mauro Silberberg\ORCID{0000-0002-2402-1100}}
\author[3]{Henning Hermjakob\ORCID{0000-0001-8479-0262}}
\author[3]{Rahuman S. Malik-Sheriff\ORCID{0000-0003-0705-9809}}
\author[1,2,$\ast$]{Hernán E. Grecco\ORCID{0000-0002-1165-4320}}

@@

% FIGURE 1
\begin{figure}[t]
  \begin{minipage}[t]{\columnwidth}    
  \begin{CodeInput}
\begin{Highlighting}[]
\KeywordTok{from} \ClassTok{poincare} \KeywordTok{import} (\ClassTok{Derivative}, \ClassTok{System},
                    \ClassTok{Variable}, \FunctionTok{initial})
\end{Highlighting}
  \end{CodeInput}
  \end{minipage}%
  \newline
  \begin{minipage}[t]{\columnwidth}
  \begin{minipage}[c]{\columnwidth}
  \begin{CodeInput}
\begin{Highlighting}[]
\KeywordTok{class} \ClassTok{Decay}\KeywordTok{(}\ClassTok{System}\KeywordTok{)}\OperatorTok{:}
  \VariableTok{x}\OperatorTok{:} \ClassTok{Variable} \OperatorTok{=} \FunctionTok{initial}\KeywordTok{(}\VariableTok{default}\OperatorTok{=}\ValueTok{1}\KeywordTok{)}
  \VariableTok{eq} \OperatorTok{=} \VariableTok{x}.\FunctionTok{derive}\KeywordTok{()} \OperatorTok{\textless{}\textless{}} \OperatorTok{-}\VariableTok{x}
\end{Highlighting}
  \end{CodeInput}
  \end{minipage}%
  %
  \newline
  \begin{minipage}[c]{0.5\columnwidth}
  \[ \frac{dx}{dt} = -x \]
  \end{minipage}%
  \begin{minipage}[c]{0.5\columnwidth}
  \[ x(0) = 1 \]
  \end{minipage}%

  \subcaption{\label{fig-first-order}First-order system of an exponential decay. The variable
  \texttt{x} stores the initial condition for \(x\), and the variable
  \texttt{eq} stores the rate equation for \(x\).}
  \end{minipage}%
  \newline
  \begin{minipage}[t]{\columnwidth}

  \begin{minipage}[c]{\columnwidth}    
  \begin{CodeInput}
\begin{Highlighting}[]
\KeywordTok{class} \ClassTok{Oscillator}\KeywordTok{(}\ClassTok{System}\KeywordTok{)}\OperatorTok{:}
  \VariableTok{x}: \ClassTok{Variable} \OperatorTok{=} \FunctionTok{initial}\KeywordTok{(\VariableTok{default}\OperatorTok{=}\ValueTok{1})}
  \VariableTok{v}: \ClassTok{Derivative} \OperatorTok{=} \VariableTok{x}.\FunctionTok{derive}\KeywordTok{(\VariableTok{initial}\OperatorTok{=}\ValueTok{0})}
  \VariableTok{eq} \OperatorTok{=} \VariableTok{v}.\FunctionTok{derive}\KeywordTok{()} \OperatorTok{\textless{}\textless{}} \OperatorTok{-}\VariableTok{x}
\end{Highlighting}
  \end{CodeInput}

  \end{minipage}%
  %
  \newline
  \begin{minipage}[c]{0.5\columnwidth}
  \[ \frac{d^2x}{dt^2} = -x \]
  \end{minipage}%
  \begin{minipage}[c]{0.5\columnwidth}
  \[
  \begin{cases}
    \begin{aligned}
      x(0) &= 1 \\
      \frac{dx}{dt}(0) &= 0
    \end{aligned}
  \end{cases}
  \]
  \end{minipage}%

  \subcaption{\label{fig-second-order}Second-order system of an harmonic oscillator, where the
  variable \texttt{v} stores the derivative \(\frac{dx}{dt}\) and the rate
  equation is specified for the derivative \texttt{v}.}
  \end{minipage}%
  \newline
  \begin{minipage}[t]{\columnwidth}
  \begin{minipage}[c]{\columnwidth}
  \begin{CodeInput}
\begin{Highlighting}[]
\KeywordTok{class} \ClassTok{BigModel}\KeywordTok{(}\ClassTok{System}\KeywordTok{)}:
  \VariableTok{x}: \ClassTok{Variable} \OperatorTok{=} \FunctionTok{initial}\KeywordTok{(\VariableTok{default}\OperatorTok{=}\ValueTok{1})}
  \VariableTok{linked} \OperatorTok{=} \ClassTok{Decay}\KeywordTok{(\VariableTok{x}\OperatorTok{=}\VariableTok{x})}
  \VariableTok{independent} \OperatorTok{=} \ClassTok{Decay}\KeywordTok{(\VariableTok{x}\OperatorTok{=}\ValueTok{2})}
\end{Highlighting}
  \end{CodeInput}
  \end{minipage}%
  %
  \newline
  \begin{minipage}[c]{0.5\columnwidth}
  \[
  \begin{cases}
    \begin{aligned}
      \frac{dx}{dt} &= -x \\
      \frac{dx_{indep}}{dt} &= -x_{indep}
    \end{aligned}
  \end{cases}
  \]
\end{minipage}%
\begin{minipage}[c]{0.5\columnwidth}
  \[
  \begin{cases}
    \begin{aligned}
      x(0) &= 1 \\
      x_{indep}(0) &= 2
    \end{aligned}
  \end{cases}
  \]
  \end{minipage}%

  \subcaption{\label{fig-composition}First-order system of two exponential decays by composition of
  the \texttt{Decay} system. The subsystem \texttt{linked} has a reference
  to the outer variable \texttt{x}, while the subsystem
  \texttt{independent} defines a new variable with initial condition $2$,
  which on the corresponding mathematical expression was named \(x_{ind}\).}
  \end{minipage}%

\end{figure}

% FIGURE 2
\begin{figure}[t]
  \includegraphics[width=\columnwidth]{ide1.pdf}
  \includegraphics[width=\columnwidth]{ide2.pdf}
  \caption{
    \label{fig-ide}
    Screenshots of Visual Studio Code showing
    tooltips (green arrows) and
    highlighted type errors (red arrows).
    On the left,
    we show that \texttt{a},
    a \texttt{Constant} assigned with \texttt{assign(...,\ constant=True)},
    can be used for \texttt{Variable b}'s initial condition.
    Instead,
    it is flagged as a type error (red underlining)
    when using \texttt{c}, a \texttt{Parameter},
    for \texttt{Variable d}'s initial condition,
    The IDE automatically recognizes \texttt{e} as an \texttt{Equation},
    and provides autocompletion of the \texttt{Model}'s components.
    A tooltip is shown when composing models (green arrow, right),
    which show the expected variables and their default values.
    The IDE highlights wrong names (\texttt{z} is not a name in \texttt{Model})
    and mismatched types (\texttt{x} is \texttt{Variable} and \texttt{a} must be a number or a \texttt{Constant})
  }  
\end{figure}

% FIGURE 3
\begin{figure}[t]
  \begin{minipage}[t]{\columnwidth}
    \begin{CodeInput}
\begin{Highlighting}[]
\KeywordTok{import} \ClassTok{numpy} \KeywordTok{as} \ClassTok{np}
\KeywordTok{from} \ClassTok{poincare} \KeywordTok{import} \ClassTok{Simulator}

\VariableTok{sim} \OperatorTok{=} \ClassTok{Simulator}\KeywordTok{(\ClassTok{Oscillator})}
\VariableTok{df} \OperatorTok{=} \VariableTok{sim}.\FunctionTok{solve}\KeywordTok{(}\VariableTok{save\_at}\OperatorTok{=}\ClassTok{np}.\FunctionTok{linspace}\KeywordTok{(}\ValueTok{0}, \ValueTok{5}, \ValueTok{101}\KeywordTok{))}
\VariableTok{df}.\FunctionTok{head}\KeywordTok{()}
\end{Highlighting}
    \end{CodeInput}
    \begin{CodeOutput}
        \begin{tabular}[]{@{}lll@{}}
            \toprule
            & x & v \\
            time & & \\
            \midrule
            0.00 & 1.000000 & 0.000000 \\
            0.05 & 0.998750 & -0.049979 \\
            0.10 & 0.995005 & -0.099835 \\
            0.15 & 0.988772 & -0.149440 \\
            0.20 & 0.980069 & -0.198672 \\
            \botrule
        \end{tabular}
    \end{CodeOutput}
  \end{minipage}%
  \hfill
  \begin{minipage}[t]{\columnwidth}
    \begin{CodeInput}
\begin{Highlighting}[]
\VariableTok{df}.\FunctionTok{plot}\KeywordTok{()}
\end{Highlighting}
    \end{CodeInput}
    %
    \begin{CodeOutput}
      \includegraphics[width=\textwidth]{cell-output}
    \end{CodeOutput}
  \end{minipage}%

  \caption{\label{fig-sim}Simulation of the \texttt{Oscillator} system
  from Figure~\ref{fig-second-order}. The output is a
  \texttt{pandas.DataFrame} with a column for each variable and the time
  as index. It is inspected and plotted with the \texttt{pandas} methods.}
\end{figure}

@@

\subsection{Extensible definition of reaction networks using SimBio}
\label{extensible-definition-of-reaction-networks-using-simbio}

% FIGURE 4
\begin{figure}[t]

  \begin{minipage}[t]{\columnwidth}
    \begin{CodeInput}
\begin{Highlighting}[]
\KeywordTok{from} \ClassTok{simbio} \KeywordTok{import} (\ClassTok{Compartment}, \ClassTok{MassAction},
                    \ClassTok{Species}, \ClassTok{RateLaw}, \FunctionTok{initial})
\end{Highlighting}
    \end{CodeInput}
  \end{minipage}%
  \newline
  \begin{minipage}[t]{\columnwidth}
    \begin{minipage}[c]{0.7\columnwidth}
      \begin{CodeInput}
\begin{Highlighting}[]
\KeywordTok{class}\ClassTok{ Model}\KeywordTok{(}\ClassTok{Compartment}\KeywordTok{)}:
  \CommentTok{"""2A {-}\textgreater{} B"""}
\VariableTok{    A}: \ClassTok{Species }\OperatorTok{=}\FunctionTok{ initial}\KeywordTok{(}\VariableTok{default}\OperatorTok{=}\ValueTok{1}\KeywordTok{)}
\VariableTok{    B}: \ClassTok{Species }\OperatorTok{=}\FunctionTok{ initial}\KeywordTok{(}\VariableTok{default}\OperatorTok{=}\ValueTok{0}\KeywordTok{)}
\VariableTok{    r }\OperatorTok{=}\ClassTok{ RateLaw}\KeywordTok{(}
\VariableTok{        reactants}\OperatorTok{=}\KeywordTok{[}\ValueTok{2} \OperatorTok{*}\VariableTok{ A}\KeywordTok{]},
\VariableTok{        products}\OperatorTok{=}\NormalTok{\KeywordTok{[}B\KeywordTok{]},}
\VariableTok{        rate}\OperatorTok{=}\ValueTok{1}\NormalTok{,}
\KeywordTok{    )}
\end{Highlighting}
      \end{CodeInput}
    \end{minipage}%
    %
    \begin{minipage}[c]{0.3\columnwidth}
      \begin{align*}
        & \begin{cases}
          A(0) = 1 \\
          B(0) = 0
          \end{cases}
        \\
        & \begin{cases}
          \frac{dA}{dt} = -2 \\
          \frac{dB}{dt} = +1
          \end{cases}
      \end{align*}
  \end{minipage}%
  \end{minipage}%
  \newline
  \begin{minipage}[t]{\columnwidth}
      \centering 
      \begin{minipage}[c]{0.7\columnwidth}
\centering 
\begin{CodeInput}
\begin{Highlighting}[]
\KeywordTok{class}\ClassTok{ Model}\KeywordTok{(}\ClassTok{Compartment}\KeywordTok{)}:
  \CommentTok{"""2A {-}\textgreater{} B"""}
  \VariableTok{A}: \ClassTok{Species} \OperatorTok{=} \FunctionTok{initial}\KeywordTok{(}\VariableTok{default}\OperatorTok{=}\ValueTok{1}\KeywordTok{)}
  \VariableTok{B}: \ClassTok{Species} \OperatorTok{=} \FunctionTok{initial}\KeywordTok{(}\VariableTok{default}\OperatorTok{=}\ValueTok{0}\KeywordTok{)}
  \VariableTok{r} \OperatorTok{=} \ClassTok{MassAction}\KeywordTok{(}
    \VariableTok{reactants}\OperatorTok{=}\KeywordTok{[}\ValueTok{2} \OperatorTok{*} \VariableTok{A}\KeywordTok{]},
    \VariableTok{products}\OperatorTok{=}\NormalTok{\KeywordTok{[}B\KeywordTok{]},}
    \VariableTok{rate}\OperatorTok{=}\ValueTok{1}\NormalTok{,}
  \KeywordTok{)}
          \end{Highlighting}
          \end{CodeInput}
      \end{minipage}%
      %
      \begin{minipage}[c]{0.3\columnwidth}
        \begin{align*}
          & \begin{cases}
              A(0) = 1 \\
              B(0) = 0
          \end{cases} \\
          & \begin{cases}
              \frac{dA}{dt} = -2 A^2 \\
              \frac{dB}{dt} = +1 A^2
            \end{cases}
        \end{align*}
      \end{minipage}%
  \end{minipage}%

  \caption{
    A reaction system for species \(A\) and \(B\)
    with initial conditions \(1\) and \(0\), respectively. A single reaction
    transforming \(2A\) into \(B\) is saved in variable \texttt{r}. The rate
    \(1\) is specified directly for \texttt{RateLaw}, and is proportional to
    the reactants for \texttt{MassAction}.
  }
  \label{fig-simbio}
\end{figure}

% FIGURE 5
\begin{figure}[t]
  \centering

  \begin{CodeInput}
\begin{Highlighting}[]
\KeywordTok{from} \ClassTok{simbio.io} \KeywordTok{import} \ClassTok{biomodels}, \ClassTok{sbml}

\CommentTok{\# from existing SBML file}
\VariableTok{model} \OperatorTok{=} \ClassTok{sbml}.\FunctionTok{load}\KeywordTok{(}\StringTok{"repressilator.sbml"}\KeywordTok{)}
\CommentTok{\# from BioModels}
\VariableTok{model} \OperatorTok{=} \ClassTok{biomodels}.\FunctionTok{load}\KeywordTok{(}\StringTok{"BIOMD12"}\KeywordTok{)}
\VariableTok{model}
\end{Highlighting}
  \end{CodeInput}

  \begin{CodeOutput}
      \begin{tabular}{@{}lll@{}}
          \multicolumn{3}{l}{Elowitz2000 - Repressilator} \\
          \toprule
          type & total & names \\
          \midrule
          variables  &  6 & PX, PY, PZ, X, Y, Z \\
          parameters & 17 & cell, beta, alpha0, alpha, eff, n, ... \\
          equations  & 12 & Reaction1, Reaction2, Reaction3, ... \\
          \botrule
      \end{tabular}
  \end{CodeOutput}

  \caption{Creation of a model from a local SBML file or one uploaded to BioModels.}
  \label{fig-simbio-io}
\end{figure}

@@
